Trustability of certification systems

### 2015-12-22 paper.tex
X.509 & PGP done

\usepackage{fancyvrb}

% some very useful LaTeX packages include:

%\usepackage{cite}      % Written by Donald Arseneau
                        % V1.6 and later of IEEEtran pre-defines the format
                        % of the cite.sty package \cite{} output to follow
                        % that of IEEE. Loading the cite package will
                        % result in citation numbers being automatically
                        % sorted and properly "ranged". i.e.,
                        % [1], [9], [2], [7], [5], [6]
                        % (without using cite.sty)
                        % will become:
                        % [1], [2], [5]--[7], [9] (using cite.sty)
                        % cite.sty's \cite will automatically add leading
                        % space, if needed. Use cite.sty's noadjust option
                        % (cite.sty V3.8 and later) if you want to turn this
                        % off. cite.sty is already installed on most LaTeX
                        % systems. The latest version can be obtained at:
                        % http://www.ctan.org/tex-archive/macros/latex/contrib/supported/cite/

\usepackage{graphicx}   % Written by David Carlisle and Sebastian Rahtz
                        % Required if you want graphics, photos, etc.
                        % graphicx.sty is already installed on most LaTeX
                        % systems. The latest version and documentation can
                        % be obtained at:
                        % http://www.ctan.org/tex-archive/macros/latex/required/graphics/
                        % Another good source of documentation is "Using
                        % Imported Graphics in LaTeX2e" by Keith Reckdahl
                        % which can be found as esplatex.ps and epslatex.pdf
                        % at: http://www.ctan.org/tex-archive/info/

%\usepackage{psfrag}    % Written by Craig Barratt, Michael C. Grant,
                        % and David Carlisle
                        % This package allows you to substitute LaTeX
                        % commands for text in imported EPS graphic files.
                        % In this way, LaTeX symbols can be placed into
                        % graphics that have been generated by other
                        % applications. You must use latex->dvips->ps2pdf
                        % workflow (not direct pdf output from pdflatex) if
                        % you wish to use this capability because it works
                        % via some PostScript tricks. Alternatively, the
                        % graphics could be processed as separate files via
                        % psfrag and dvips, then converted to PDF for
                        % inclusion in the main file which uses pdflatex.
                        % Docs are in "The PSfrag System" by Michael C. Grant
                        % and David Carlisle. There is also some information
                        % about using psfrag in "Using Imported Graphics in
                        % LaTeX2e" by Keith Reckdahl which documents the
                        % graphicx package (see above). The psfrag package
                        % and documentation can be obtained at:
                        % http://www.ctan.org/tex-archive/macros/latex/contrib/supported/psfrag/

%\usepackage{subfigure} % Written by Steven Douglas Cochran
                        % This package makes it easy to put subfigures
                        % in your figures. i.e., "figure 1a and 1b"
                        % Docs are in "Using Imported Graphics in LaTeX2e"
                        % by Keith Reckdahl which also documents the graphicx
                        % package (see above). subfigure.sty is already
                        % installed on most LaTeX systems. The latest version
                        % and documentation can be obtained at:
                        % http://www.ctan.org/tex-archive/macros/latex/contrib/supported/subfigure/

\usepackage{url}        % Written by Donald Arseneau
                        % Provides better support for handling and breaking
                        % URLs. url.sty is already installed on most LaTeX
                        % systems. The latest version can be obtained at:
                        % http://www.ctan.org/tex-archive/macros/latex/contrib/other/misc/
                        % Read the url.sty source comments for usage information.

%\usepackage{stfloats}  % Written by Sigitas Tolusis
                        % Gives LaTeX2e the ability to do double column
                        % floats at the bottom of the page as well as the top.
                        % (e.g., "\begin{figure*}[!b]" is not normally
                        % possible in LaTeX2e). This is an invasive package
                        % which rewrites many portions of the LaTeX2e output
                        % routines. It may not work with other packages that
                        % modify the LaTeX2e output routine and/or with other
                        % versions of LaTeX. The latest version and
                        % documentation can be obtained at:
                        % http://www.ctan.org/tex-archive/macros/latex/contrib/supported/sttools/
                        % Documentation is contained in the stfloats.sty
                        % comments as well as in the presfull.pdf file.
                        % Do not use the stfloats baselinefloat ability as
                        % IEEE does not allow \baselineskip to stretch.
                        % Authors submitting work to the IEEE should note
                        % that IEEE rarely uses double column equations and
                        % that authors should try to avoid such use.
                        % Do not be tempted to use the cuted.sty or
                        % midfloat.sty package (by the same author) as IEEE
                        % does not format its papers in such ways.

\usepackage{amsmath}    % From the American Mathematical Society
                        % A popular package that provides many helpful commands
                        % for dealing with mathematics. Note that the AMSmath
                        % package sets \interdisplaylinepenalty to 10000 thus
                        % preventing page breaks from occurring within multiline
                        % equations. Use:
%\interdisplaylinepenalty=2500
                        % after loading amsmath to restore such page breaks
                        % as IEEEtran.cls normally does. amsmath.sty is already
                        % installed on most LaTeX systems. The latest version
                        % and documentation can be obtained at:
                        % http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/

% Other popular packages for formatting tables and equations include:

%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty which improves the
% LaTeX2e array and tabular environments to provide better appearances and
% additional user controls. array.sty is already installed on most systems.
% The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/tools/

% V1.6 of IEEEtran contains the IEEEeqnarray family of commands that can
% be used to generate multiline equations as well as matrices, tables, etc.

% Also of notable interest:
% Scott Pakin's eqparbox package for creating (automatically sized) equal
% width boxes. Available:
% http://www.ctan.org/tex-archive/macros/latex/contrib/supported/eqparbox/

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.

@@

% Define document title and author
	\title{Trustability of certification systems}
	\author{Alexandre Kervadec
	\thanks{Tutor : A.Guermouche}}
	\markboth{University of Bordeaux - Master 2 Computing Sciences}{}
	\maketitle

% Write abstract here
\begin{abstract}
This paper is a synthesis of a E.Gerck's paper : \textit{"Overview of Certification systems: X.509, CA, PGP and SKIP"}\cite{gerck1998overview}, but not only. It first gives an overview of (the mains) certification systems which are X.509 and CAs, PGP, SKIP, DANE and the Certificate transparency by Google, with thoughts on pros and cons of each system, on a technical point of view and about the government's stranglehold on data exchanges.
\end{abstract}

@@

First sections talk about existing (sections \ref{x509} and \ref{pgp}) or extinguished (section \ref{skip}) certification systems. Then, further sections (\ref{dane} and \ref{certtrans}) deals with new, not implemented yet, technologies of certification that may answer to current issues.

@@

X.509 and CAs\cite{rfc3647} infrastructure is based on a directory method.
With this kind of certification system, there are three different entities, which are :

\begin{enumerate}
	\item \textbf{CA : Certification Authority}, an entity at controls the authentication services and the management of certificates. It can be public (like banks with their clients), commercial (like Verisign which sells its services) or private (like a compagny department, for an internal purpose).
	\item \textbf{Subscriber :} the entity which sends informations to the CA to add it to his certificate. The entity is one that need to be trusted by the next entity (\textit{user}).
	\item \textbf{User :} ask infos to CA(s), it's central to the process, since the user party is relying on the informations and is thus at risk.
\end{enumerate}

@@

\begin{itemize}
	\item The content of a certificate needs to be discussed, as well as certificate revocation
	\item Issues about DN (Distinguished Name) and CA : a user can possess one or more DN, and use one or more DN on one or more DN
	\item Validation of the user, using an ID, which is easily subject to fraud
\end{itemize}

@@

Some kind of conclusion about harmonization (lack of), in a world wide vision.

@@

PGP, created thanks to Phil Zimmermann researches.

[TODO : Schema]

It is based on the \textit{web-of-trust}, if the client doesn't know a CA which gave him a certificate, he ask to the PGP ring.
In that ring, even if the client doesn't know everyone, someone will know him, thus the web-of-trust principle.
Each member of the ring have a trust evaluation of CAs. If the client knows someone he trust that trust the CA he doesn't know, then he accept the certificate and trust the CA.
But if the client trust someone who doesn't trust the CA, he reject the CA.
The big point is when nobody knows the CA, and when trusted friends of the client have a medium evaluation of the CA : when to trust, when to reject, at which level of trust?

@@

\subsection{Cetificate Transparency}
\label{certtrans}

@@

\bibitem{gerck1998overview} E.Gerck, Overview of Certification Systems: X. 509, CA, PGP and SKIP, 1998. http://www.blackhat.com/presentations/bh-usa-99/EdGerck/certover.pdf

@@

\bibitem{skip2010user} Oracle, SunScreen SKIP User's Guide, Release 1.1, 2010. http://docs.oracle.com/cd/E19957-01/805-5743/6j5dvnrfs/index.html

@@

\bibitem{certtransp2013} Laurie, B., Langley, A., \& Kasper, E. (2013). Certificate transparency (No. RFC 6962). https://www.rfc-editor.org/rfc/pdfrfc/rfc6962.txt.pdf

\bibitem{rfc3647} Solo, D., Housley, R., \& Ford, W. (1999). Internet X. 509 public key infrastructure certificate and CRL profile. http://tools.ietf.org/html/rfc2459

>>>>

\usepackage{fancyvrb}
\usepackage{graphicx}
\usepackage{url}
\usepackage{amsmath}
\usepackage{tikz}
\usepackage{verbatim}

@@

% Define document title and author
	\title{Trustability of certification systems}
	\author{Alexandre Kervadec
	\thanks{Tutor : A.Guermouche}}
	\markboth{University of Bordeaux -- Master 2 Computing Sciences - Networking and mobile systems}{}
	\maketitle

% Write abstract here
\begin{abstract}
This paper is a synthesis of a E.Gerck's paper : \textit{"Overview of Certification systems: X.509, CA, PGP and SKIP"}\cite{gerck1998overview}, but not only. First, it gives an overview of (the mains) certification systems which are X.509 and CAs, PGP, SKIP, DANE and the Certificate transparency by Google, with thoughts on pros and cons of each system, on a technical point of view and about the government's stranglehold on data exchanges.
\end{abstract}

@@

First sections talk about existing (sections \ref{x509} and \ref{pgp}) or extinguished (section \ref{skip}) certification systems. Then, further sections (\ref{dane} and \ref{certtrans}) deals with new, not implemented yet, technologies of digital certification which may solve current issues.

@@

X.509 and CAs\cite{rfc3647} infrastructure is based on a directory method.\\
With this kind of certification system, there are three different entities, which are :

\begin{enumerate}
	\item \textbf{CA : Certification Authority}, an entity that controls authentication services and management of digital certificates. It can be public (like banks with their clients), commercial (such as \textit{Verisign} which sells its services) or private (like a compagny department, for an internal purposes).
	\item \textbf{Subscriber :} the entity which sends informations to the CA to add it to his certificate. This entity is one that need to be trusted by the \textit{user} entity.
	\item \textbf{User :} ask informations about \textit{subscribers} to CA(s), it's central to the process, since the user party is relying on the informations given by \textit{CAs} and is thus at risk.
\end{enumerate}

The way that certification system is the following:
\begin{itemize}
	\item A subscriber gives informations to the CA, in order to create a digital certificate for itself in the CA directory.
	\item The CA receive those informations and put it in a certificate.
	\item A user (the public in general) receive a certificate from a subscriber. In order to authenticate this subscriber, the client ask to the CA if the transmitted certificate matches with the subscriber's identity.
	\item The CA answer to the user to give him the precious informations about the subscriber's identity.
\end{itemize}

There is a not outwardly perceived entity in this system, it is the Naming Authority (DN). It provides to the CA a unique distinguished name (DN) which matches to the subscriber identity in the certificate.\\
Note that the CA can double in NA, but they provide two different services: the CA certificate refers to a name but does not denote it, the NA does.\\

@@

\begin{itemize}
	\item The content of a certificate needs to be discussed, as well as certificate revocation. For example, a subscriber can generate multiple DN for a same CA or the same DN to multiple CAs, so maybe, a better subscriber identificator need to be added.
	\item Are the validation procedures for the certified data that is included in a certificate solid enough? Because each CA has its own self-defined rules (Certification Practice Statement\footnote{CPS}), which can by completly different from one CA to another.
\end{itemize}

@@

Also, regarrding to the law, CAs deny all leak of information from themselves.

Some kind of conclusion about harmonization (lack of), in a world wide vision, if space there is.

@@

PGP, created thanks to Phil Zimmermann researches. It has two parts: certification and encryption, following text only deals with certification.\\
Compared to X.509, PGP is more \textit{internet-like}, this because of its \textit{introducer-model} base.\\
PGP depends on a chain of authentificators: the users themselves. This chain forms a ring, but not in a closed way but in a mathematical way like a list or a \textit{web-of-trust}. You may not know everyone in this ring, but you can assume that you will know somebody who know this user. You can also assume that different rings can have some contact points to garantee the referrals.

\begin{figure}[!hbt]
	\begin{center}

		\begin{tikzpicture}[shorten >=1pt,->]
			\tikzstyle{vertex}=[circle,fill=black!25,minimum size=25pt,inner sep=0pt]

			\foreach \name/\angle/\text in {P-1/234/Alice, P-2/162/Bob, 
                                  P-3/90/Oscar, P-4/18/Jack, P-5/-54/Alex}
				\node[vertex,xshift=6cm,yshift=.5cm] (\name) at (\angle:2cm) {$\text$};

			\draw (P-1) -- (P-5);
			\draw[dashed,->] (P-2) -- (P-4);
			\draw (P-4) -- (P-3);
			\draw (P-3) -- (P-1);
			\draw (P-1) -- (P-2);
			\draw (P-2) -- (P-1);
			\draw (P-3) -- (P-5);
			\draw (P-5) -- (P-3);

		\end{tikzpicture}
		\caption{Communities of trust: Jack introduces Oscar to Bob's public-key certificate before Bob receives it.}
		\label{fig:web-of-trust}
	\end{center}
\end{figure}

Thus, we can figure out that PGP breaks the traditional hierarchical trust architecture with this \textit{web-of-trust} approach.\\

Let's get an example\footnote{Example from ''The PGP trust model''\cite{caronni2000}}: Bob want to exchange datas with Oscar, which he didn't knew before. But Jack, a colleague of Bob's, signs Oscar's public-key certificate whch he knows is authentic.\\
Oscar then forwards his signed certificate to Bob who wishes to communicate with Oscar privately.\\
Bob, who knows and truts Jack as an \textit{introducer}, find out, after verification, that Jack is among Oscar's certificate signer. Therefor, Bob can be confident that Oscar's public key is authentic.

However, had Bob not known or trusted Oscar's signers, including Jack, he would have been skeptical about the authenticity of Oscar's public-key. Oscar would have to find another introducer whom Bob trusts to sign Oscar's public-key certificate. This is illustrated in Figure \ref{fig:web-of-trust}.\\
This is illustrated in Figure 1, where the dotted arrow indicates ''Bob trusts Jack as introducer'' and the solid arrow indicates ''Jack trusts Oscar’s public key validity''.

PGP also includes a public-key certificate and an introducer trustwhorthiness. They have different levels, for a public-key certificate they are:
\begin{itemize}
	\item \textit{Undefined:} we cannot say wether this public-key is valid or not.
	\item \textit{Marginal:} this public-key \textit{may} be valid but we cannot be sure.
	\item \textit{Complete:} we can be fully confident that this public-key is valid.
\end{itemize}
For an introducer, those levels of trustworthiness are the following:
\begin{itemize}
	\item \textit{Full:} this public-key is fully trusted to introduce another public-key.
	\item \textit{Marginal:} this public-key can be trusted to introduce another public-key, but, it is uncertain whether it is fully competent to do that.
	\item \textit{Untrustworthy:} this public-key should not be trustedto introduce another, therefore any occurence of this key as a signature on another public-key should be ignored.
\end{itemize}

Only the trustworthiness of public-key's validity is automatically evaluated by PGP. Introducer's worthiness is manually assigned by each user of to the public key, and exists only within each individual user's public-key ring\footnote{PGP allows users to have files representing multipke \textit{key rings} to store public or secret keys\cite{caronni2000}.}.

The big point is the use of such a tool in a world wide range: that is raisonnably not possible. We can assume to use it in a little group, where trustworthiness is effective, but in a more wide area, it is problematic to trust such a amount of people, or find trusted introducers.

@@

\subsection{CATA: Certificate Authority Transparency and Auditability}
\label{certtrans}

@@

\bibitem{gerck1998overview} E.Gerck, Overview of Certification Systems: X. 509, CA, PGP and SKIP, 1998.

@@

\bibitem{skip2010user} Oracle, SunScreen SKIP User's Guide, Release 1.1, 2010. %http://docs.oracle.com/cd/E19957-01/805-5743/6j5dvnrfs/index.html

\bibitem{caronni2000} Caronni, G. (2000). Walking the web of trust. In Enabling Technologies: Infrastructure for Collaborative Enterprises, 2000.(WET ICE 2000). Proeedings. IEEE 9th International Workshops on (pp. 153-158). IEEE. %http://system.khl.mydns.jp/college/WOT/The_PGP_Trust_Model.pdf

@@

\bibitem{certtransp2013} Laurie, B., Langley, A., \& Kasper, E. (2013). Certificate transparency (No. RFC 6962). %https://www.rfc-editor.org/rfc/pdfrfc/rfc6962.txt.pdf

\bibitem{rfc3647} Solo, D., Housley, R., \& Ford, W. (1999). Internet X. 509 public key infrastructure certificate and CRL profile. %http://tools.ietf.org/html/rfc2459
